---
title: "data_cleaning"
output: html_document

---
Load packages
```{r, include = F}
gc()
library(tidyverse) #for data work
library(data.table) #for data loading

#watch these packages for dummies
install.packages("fastDummies") # --- recipes package might also work or tidymodels? (step_dummy maybe work darn recipe use confusion) (or not- different method) - issue: combo 2 years - check
library(fastDummies)
```

Download the data
```{r, include = F}
gc()

pull1_csv <- fread("C:\\Users\\njrich\\Downloads\\pull1.csv")

#note might need to redownload to factor in ASECWT weights - watch individual v household serial and rules of weights - questions about ASECWT only in IPUMS CPS maybe also widow var... hmm.....
```

Select out unneeded variables
```{r, include = F}
gc()
pull1_csv <- pull1_csv %>%
  select(-SAMPLE, -CBSERIAL, -HHWT, -CLUSTER, -STRATA, -GQ)

#note: need to add ASECWT weight var after looking

```

Identify individuals in same-sex relationships  
```{r, include = F}
gc()
partner <- pull1_csv %>%
  select(YEAR, SERIAL, PERNUM, SEX) 

gc()
df_partners <- pull1_csv %>%
  left_join(partner, join_by(YEAR, SERIAL, SPLOC == PERNUM), suffix = c(".indv", ".partner"))

gc()
df_samesex <- df_partners %>% #WATCH CLARITY WHERE SPLOC IS
  mutate(in_samesex = case_when(
    (SEX.indv == 1) & (SEX.partner == 1) ~ 1, #gay
    (SEX.indv == 2) & (SEX.partner == 2) ~ 1, #gay
    (SEX.indv == 2) & (SEX.partner == 1) ~ 0, #straight
    (SEX.indv == 1) & (SEX.partner == 2) ~ 0, #straight
    (SEX.indv == 9) | (SEX.partner == 9) ~ NA #NA 
  ))

rm(partner, df_partners, pull1_csv)
#check all mutate works             
```

Label states/input legalization year/length of time since legalization (use FIP) 

PROBLEM: MISSING YEARS AND STATE NAMES WITH CHUNKING
ChatGPT solution doesn't seem to work... keep thinking with everything else return to og and come back to later
```{r, include = F}

#label states
gc()
df_samesex <- df_samesex %>%
  mutate(state_name = NA) %>% #ok this seems to work
  mutate(state_name = case_when( 
    STATEFIP == 1 ~ "AL",
    STATEFIP == 2 ~ "AK",
    STATEFIP == 4 ~ "AZ",
    STATEFIP == 5 ~ "AR",
    STATEFIP == 6 ~ "CA",
    STATEFIP == 8 ~ "CO",
    STATEFIP == 9 ~ "CT",
    STATEFIP == 10 ~ "DE"))

gc()
df_samesex <- df_samesex %>%
  mutate(state_name = case_when(
    STATEFIP == 11 ~ "DC",
    STATEFIP == 12 ~ "FL",
    STATEFIP == 13 ~ "GA",
    STATEFIP == 15 ~ "HI",
    STATEFIP == 16 ~ "ID",
    STATEFIP == 17 ~ "IL",
    STATEFIP == 18 ~ "IN",
    STATEFIP == 19 ~ "IA",
    STATEFIP == 20 ~ "KS")) 

gc()
df_samesex <- df_samesex %>%
  mutate(state_name = case_when(
    STATEFIP == 21 ~ "KY",
    STATEFIP == 22 ~ "LA",
    STATEFIP == 23 ~ "MI",
    STATEFIP == 24 ~ "MD",
    STATEFIP == 25 ~ "MA",
    STATEFIP == 26 ~ "ME",
    STATEFIP == 27 ~ "MN",
    STATEFIP == 28 ~ "MS",
    STATEFIP == 29 ~ "MO",
    STATEFIP == 30 ~ "MT")) 

gc()
df_samesex <- df_samesex %>%
  mutate(state_name = case_when(
    STATEFIP == 31 ~ "NE",
    STATEFIP == 32 ~ "NV",
    STATEFIP == 33 ~ "NH",
    STATEFIP == 34 ~ "NJ",
    STATEFIP == 35 ~ "NM",
    STATEFIP == 36 ~ "NY",
    STATEFIP == 37 ~ "NC",
    STATEFIP == 38 ~ "ND",
    STATEFIP == 39 ~ "OH",
    STATEFIP == 40 ~ "OK")) 

gc()
df_samesex <- df_samesex %>% #this seems to be the only clause that works...
  mutate(state_name = case_when(
    STATEFIP == 41 ~ "OR",
    STATEFIP == 42 ~ "PA",
    STATEFIP == 44 ~ "RI",
    STATEFIP == 45 ~ "SC",
    STATEFIP == 46 ~ "SD",
    STATEFIP == 47 ~ "TN",
    STATEFIP == 48 ~ "TX",
    STATEFIP == 49 ~ "UT",
    STATEFIP == 50 ~ "VT",
    STATEFIP == 51 ~ "VA",
    STATEFIP == 53 ~ "WA",
    STATEFIP == 54 ~ "WV",
    STATEFIP == 55 ~ "WI",
    STATEFIP == 56 ~ "WY",
    STATEFIP == 72 ~ "PR",
    STATEFIP == 97 ~ "OMI", #Overseas Military Installations
    STATEFIP == 99 ~ NA #redundant now
  )) 


##TEMP SOLN
df_samesex <- df_samesex %>%
  filter(!is.na(state_name)) #TEMPORARY SOLUTION HMM
##TEMP SOLN


#label legalization date from Gerstmann 2012 OG pqper (see issues)
gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = NA) %>%
  mutate(date_legalization = case_when ( 
    state_name == "AL" ~ 2015,
    state_name == "AK" ~ 2014,
    state_name == "AZ" ~ 2014,
    state_name == "AR" ~ 2015,
    state_name == "CA" ~ 2013, #watch leglized/repealed issue
    state_name == "CO" ~ 2014,
    state_name == "CT" ~ 2008,
    state_name == "DE" ~ 2013,
    state_name == "DC" ~ 2009,
    state_name == "FL" ~ 2015)) 

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when(
    state_name == "GA" ~ 2015,
    state_name == "HI" ~ 2013,
    state_name == "ID" ~ 2014,
    state_name == "IL" ~ 2013,
    state_name == "IN" ~ 2014,
    state_name == "IA" ~ 2009,
    state_name == "KS" ~ 2014,
    state_name == "KY" ~ 2015,
    state_name == "LA" ~ 2015,
    state_name == "ME" ~ 2012)) 

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when(
    state_name == "MD" ~ 2012,
    state_name == "MA" ~ 2003,
    state_name == "MI" ~ 2015,
    state_name == "MN" ~ 2013,
    state_name == "MS" ~ 2015,
    state_name == "MO" ~ 2015,
    state_name == "MT" ~ 2014,
    state_name == "NE" ~ 2015,
    state_name == "NV" ~ 2014,
    state_name == "NH" ~ 2009)) 

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when(
    state_name == "NJ" ~ 2013,
    state_name == "NM" ~ 2013,
    state_name == "NY" ~ 2011,
    state_name == "NC" ~ 2014,
    state_name == "ND" ~ 2015,
    state_name == "OH" ~ 2015,
    state_name == "OK" ~ 2014,
    state_name == "OR" ~ 2014,
    state_name == "PA" ~ 2014,
    state_name == "RI" ~ 2013)) 

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when(
    state_name == "SC" ~ 2014,
    state_name == "SD" ~ 2015,
    state_name == "TN" ~ 2015,
    state_name == "TX" ~ 2015,
    state_name == "UT" ~ 2014,
    state_name == "VT" ~ 2009,
    state_name == "VA" ~ 2014,
    state_name == "WA" ~ 2012,
    state_name == "WV" ~ 2014,
    state_name == "WI" ~ 2014,
    state_name == "WY" ~ 2014,
    state_name == "PR" ~ 2015, #from googling
    state_name == "OMI" ~ 2015 #from guessing
  ))

#make time_since variable
gc()
df_samesex <- df_samesex %>%
  mutate(time_since_legalization = if_else(YEAR - date_legalization < 0, NA, YEAR - date_legalization)) #check if this makes sense

#save output of this somewhere for ease of use?

```

Make pct same-sex by state and year var (NOTE: WILL NEED TO ADD WEIGHTS EVENTUALLY -I believe only here tho)
```{r, include = F}
gc()
samesex_staterate <- df_samesex %>%
  filter(!is.na(MIGPLAC1)) %>% #get interstate migrants only
  group_by(YEAR, state_name) %>%
  summarize(samesex_state_count = sum(in_samesex, na.rm = T)) %>%
  ungroup() %>%
  group_by(YEAR) %>%
  mutate(year_totals = sum(samesex_state_count)) %>%
  ungroup() %>%
  mutate(staterate_year = samesex_state_count/year_totals) %>%
  select(YEAR, state_name, staterate_year)

#watch NA state names taking large percent of migrants hmm check out (OPE WE STILL HAVE MISSING STATE NAMES OI), FIGURE OUT WAY TO KEEP OTHER COLUMNS INDIVIDUAL LEVEL DATA TOO HMM YAY HETEROGENEITY ANALYSIS PREP... more merges? oi., drop NAs

```
PROBLEM: MISSING YEARS AND STATE NAMES WITH CHUNKING

Make dummies for analysis based on time_since_legalization (other methods possible with different packages?)
```{r, include = F}
gc()
?model_matrix #concerned... actually make matrix

#other options
#if_else method
#fastDummies library + dummy_cols()

```
Other data cleaning - watch weights see above for right variable - watch NAs too

Initial Regression