---
title: "data_cleaning"
output: html_document

---
Load packages
```{r, include = F}
gc()
install.packages("dataCompareR")
library(dataCompareR) #for rCompare


library(data.table) #for data loading
library(haven) #for writing to dta
library(readxl) #for reading in FRED data
library(tidyverse) #for data work
```

Download IPUMS data
```{r, include = F}
gc()

pull1_csv <- fread("C:\\Users\\njrich\\Downloads\\pull1.csv")

```

Select out unneeded variables
```{r, include = F}
gc()
pull1_csv <- pull1_csv %>%
  select(-SAMPLE, -CBSERIAL, -HHWT, -CLUSTER, -STRATA, -GQ) %>%
  filter(SPLOC != 0) #filter out people not in relationships

```

Identify individuals in same-sex relationships (with weights)  
```{r, include = F}
gc()
partner <- pull1_csv %>%
  select(YEAR, SERIAL, PERNUM, SEX) 

gc()
df_partners <- pull1_csv %>%
  left_join(partner, join_by(YEAR, SERIAL, SPLOC == PERNUM), suffix = c(".indv", ".partner"))

gc()
df_samesex <- df_partners %>% 
  mutate(in_samesex = case_when(
    (SEX.indv == 1) & (SEX.partner == 1) ~ 1, #gay
    (SEX.indv == 2) & (SEX.partner == 2) ~ 1, #gay
    (SEX.indv == 2) & (SEX.partner == 1) ~ 0, #straight
    (SEX.indv == 1) & (SEX.partner == 2) ~ 0, #straight
    (SEX.indv == 9) | (SEX.partner == 9) ~ NA #NA 
  )) %>%
  mutate(in_samesex_weighted = in_samesex * PERWT) #NOTE: justification on weighting methodology: https://forum.ipums.org/t/identifying-same-sex-and-opposite-sex-couples/3878 for more help; https://forum.ipums.org/t/how-to-identify-same-sex-couples/2554/5

rm(partner, df_partners, pull1_csv)


#next steps: download attached partner information directly from IPUMS (see pull3)
```

Label states/input legalization year/length of time since legalization (use FIP) 

```{r, include = F}

#label states using chunking for memory management
gc()
df_samesex <- df_samesex %>%
  mutate(state_name = as.character(STATEFIP)) 

df_samesex <- df_samesex %>% 
  mutate(state_name = case_when( 
    state_name == "1" ~ "AL",
    state_name == "2" ~ "AK",
    state_name == "4" ~ "AZ",
    state_name == "5" ~ "AR",
    state_name == "6" ~ "CA",
    state_name == "8" ~ "CO",
    state_name == "9" ~ "CT",
    state_name == "10" ~ "DE",
    TRUE ~ state_name))

gc()
df_samesex <- df_samesex %>%
  mutate(state_name = case_when(
    state_name == "11" ~ "DC",
    state_name == "12" ~ "FL",
    state_name == "13" ~ "GA",
    state_name == "15" ~ "HI",
    state_name == "16" ~ "ID",
    state_name == "17" ~ "IL",
    state_name == "18" ~ "IN",
    state_name == "19" ~ "IA",
    state_name == "20" ~ "KS",
    TRUE ~ state_name)) 

gc()
df_samesex <- df_samesex %>%
  mutate(state_name = case_when(
    state_name == "21" ~ "KY",
    state_name == "22" ~ "LA",
    state_name == "23" ~ "MI",
    state_name == "24" ~ "MD",
    state_name == "25" ~ "MA",
    state_name == "26" ~ "ME",
    state_name == "27" ~ "MN",
    state_name == "28" ~ "MS",
    state_name == "29" ~ "MO",
    state_name == "30" ~ "MT",
    TRUE ~ state_name)) 

gc()
df_samesex <- df_samesex %>%
  mutate(state_name = case_when(
    state_name == "31" ~ "NE",
    state_name == "32" ~ "NV",
    state_name == "33" ~ "NH",
    state_name == "34" ~ "NJ",
    state_name == "35" ~ "NM",
    state_name == "36" ~ "NY",
    state_name == "37" ~ "NC",
    state_name == "38" ~ "ND",
    state_name == "39" ~ "OH",
    state_name == "40" ~ "OK",
    TRUE ~ state_name)) 

gc()
df_samesex <- df_samesex %>% 
  mutate(state_name = case_when(
    state_name == "41" ~ "OR",
    state_name == "42" ~ "PA",
    state_name == "44" ~ "RI",
    state_name == "45" ~ "SC",
    state_name == "46" ~ "SD",
    state_name == "47" ~ "TN",
    state_name == "48" ~ "TX",
    state_name == "49" ~ "UT",
    state_name == "50" ~ "VT",
    state_name == "51" ~ "VA",
    state_name == "53" ~ "WA",
    state_name == "54" ~ "WV",
    state_name == "55" ~ "WI",
    state_name == "56" ~ "WY",
    state_name == "72" ~ "PR",
    state_name == "97" ~ "OMI", #Overseas Military Installations
    state_name == "99" ~ NA,
    TRUE ~ state_name 
  )) 



#label legalization date from Gerstmann "2012" pqper
gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = state_name) 

df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when ( 
    date_legalization == "AL" ~ "2015",
    date_legalization == "AK" ~ "2014",
    date_legalization == "AZ" ~ "2014",
    date_legalization == "AR" ~ "2015",
    date_legalization == "CA" ~ "2013", #watch legalized/repealed timeline
    date_legalization == "CO" ~ "2014",
    date_legalization == "CT" ~ "2008",
    date_legalization == "DE" ~ "2013",
    date_legalization == "DC" ~ "2009",
    date_legalization == "FL" ~ "2015",
    TRUE ~ date_legalization)) 

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when(
    date_legalization == "GA" ~ "2015",
    date_legalization == "HI" ~ "2013",
    date_legalization == "ID" ~ "2014",
    date_legalization == "IL" ~ "2013",
    date_legalization == "IN" ~ "2014",
    date_legalization == "IA" ~ "2009",
    date_legalization == "KS" ~ "2014",
    date_legalization == "KY" ~ "2015",
    date_legalization == "LA" ~ "2015",
    date_legalization == "ME" ~ "2012",
    TRUE ~ date_legalization)) 

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when(
    date_legalization == "MD" ~ "2012",
    date_legalization == "MA" ~ "2003",
    date_legalization == "MI" ~ "2015",
    date_legalization == "MN" ~ "2013",
    date_legalization == "MS" ~ "2015",
    date_legalization == "MO" ~ "2015",
    date_legalization == "MT" ~ "2014",
    date_legalization == "NE" ~ "2015",
    date_legalization == "NV" ~ "2014",
    date_legalization == "NH" ~ "2009",
    TRUE ~ date_legalization)) 

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when(
    date_legalization == "NJ" ~ "2013",
    date_legalization == "NM" ~ "2013",
    date_legalization == "NY" ~ "2011",
    date_legalization == "NC" ~ "2014",
    date_legalization == "ND" ~ "2015",
    date_legalization == "OH" ~ "2015",
    date_legalization == "OK" ~ "2014",
    date_legalization == "OR" ~ "2014",
    date_legalization == "PA" ~ "2014",
    date_legalization == "RI" ~ "2013",
    TRUE ~ date_legalization)) 

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = case_when(
    date_legalization == "SC" ~ "2014",
    date_legalization == "SD" ~ "2015",
    date_legalization == "TN" ~ "2015",
    date_legalization == "TX" ~ "2015",
    date_legalization == "UT" ~ "2014",
    date_legalization == "VT" ~ "2009",
    date_legalization == "VA" ~ "2014",
    date_legalization == "WA" ~ "2012",
    date_legalization == "WV" ~ "2014",
    date_legalization == "WI" ~ "2014",
    date_legalization == "WY" ~ "2014",
    date_legalization == "PR" ~ "2015", 
    date_legalization == "OMI" ~ "2015", 
    TRUE ~ date_legalization 
  ))

gc()
df_samesex <- df_samesex %>%
  mutate(date_legalization = as.numeric(date_legalization))

#make time_since variable
gc()
df_samesex <- df_samesex %>%
  mutate(time_since_legalization = if_else(((YEAR - date_legalization) < 0), NA, (YEAR - date_legalization))) 

#make time variable
gc()
df_samesex <- df_samesex %>%
  mutate(time = YEAR - 2001) %>%
  mutate(timesq = time ^2)

```

Export partially labeled data
```{r, include = F}
gc()
write.csv(df_samesex, "C:\\Users\\njrich\\Desktop\\same-sex-migration\\raw_labelled.csv")

```


Create state pop table for pop weighting
```{r, include = F}
gc()

#read in data
#FRED <- read_excel("/Users/njrich/Desktop/Econ495/same-sex-migration/Data/FRED_statepops/STATEPOP.xls", sheet = "Annual") #for Mac
FRED <- read_excel("C:\\Users\\njrich\\Downloads\\STATEPOP.xls", sheet = "Annual") #for Windows

#match up columns
FRED <- FRED %>%
  pivot_longer(AKPOP:WYPOP, names_to = "state_name") %>%
  rename("state_pop" = "value") %>%
  rename("YEAR" = "DATE")

#match up strings
FRED <- FRED %>%
  mutate(state_name = str_remove(state_name, "POP")) %>%
  mutate(YEAR = as.character(YEAR)) %>%
  mutate(YEAR = str_remove(YEAR, "-01-01")) %>%
  mutate(YEAR = as.numeric(YEAR)) %>%
  mutate(state_pop = as.character(state_pop)) %>%
  mutate(state_pop = str_remove(state_pop, "\\.")) %>%
  mutate(state_pop = as.numeric(state_pop)) %>%
  mutate(time = YEAR - 2001) %>% 
  mutate(timesq = time ^2)

```

##MARCEN REPLICATION

Filter data to match Marcén/Morales target population
```{r, include = F}
gc()

df_samesex_mm <- df_samesex %>%
  filter(in_samesex == 1) %>% 
  filter(AGE >= 30 & AGE <= 64) %>%
  filter(YEAR >= 2001 & YEAR <= 2015) %>% 
  filter(MARST == 4 | MARST == 5 | MARST == 6) #includes those divorced, widowed, never married/single (NOTE: this excludes those married, separated, blank/missing)

#filter by sample
df_samesex_mm_gen <- df_samesex_mm #general sample
df_samesex_mm_m <- df_samesex_mm %>% #men only
                    filter(SEX.indv == 1)
df_samesex_mm_f <- df_samesex_mm %>% #women only
                    filter(SEX.indv == 2)

#next steps: adjust filters to exclude individuals mis-identified as partners by IPUMS
```

##EXTENSION - 2001 - 2019
```{r, include = F}
gc()

df_samesex_2019 <- df_samesex %>%
  filter(in_samesex == 1) %>% 
  filter(AGE >= 30 & AGE <= 64) %>%
  filter(YEAR >= 2001 & YEAR <= 2019) %>% 
  filter(MARST == 4 | MARST == 5 | MARST == 6) #includes those divorced, widowed, never married/single (NOTE: this excludes those married, separated, blank/missing)

#filter by sample
df_samesex_2019_gen <- df_samesex_2019 #general sample
df_samesex_2019_m <- df_samesex_2019 %>% #men only
                    filter(SEX.indv == 1)
df_samesex_2019_f <- df_samesex_2019 %>% #women only
                    filter(SEX.indv == 2)

```

##EXTENSION - 2001 - 2015 by income (USE OFFICIAL METRICS IN THE FUTURE SAMPLES TOO SMALL COME BACK LATER
```{r, include = F}
gc()

df_samesex_inc <- df_samesex %>%
  filter(in_samesex == 1) %>% 
  filter(AGE >= 30 & AGE <= 64) %>%
  filter(YEAR >= 2001 & YEAR <= 2015) %>% 
  filter(MARST == 4 | MARST == 5 | MARST == 6) #includes those divorced, widowed, never married/single (NOTE: this excludes those married, separated, blank/missing)


df_samesex_inc <- df_samesex_inc %>%
  mutate(inc_class_year = case_when( #data in current dollars from IRS, below 50th percentile = 1, higher income = 2
    YEAR == 2001 & INCTOT < 31418 ~ 1,
    YEAR == 2002 & INCTOT < 31299 ~ 1,
    YEAR == 2003 & INCTOT < 31447 ~ 1,
    YEAR == 2004 & INCTOT < 32622 ~ 1,
    YEAR == 2005 & INCTOT < 33484 ~ 1,
    YEAR == 2006 & INCTOT < 34417 ~ 1,
    YEAR == 2007 & INCTOT < 35541 ~ 1,
    YEAR == 2008 & INCTOT < 35340 ~ 1,
    YEAR == 2009 & INCTOT < 34156 ~ 1,
    YEAR == 2010 & INCTOT < 34338 ~ 1,
    YEAR == 2011 & INCTOT < 34823 ~ 1,
    YEAR == 2012 & INCTOT < 36055 ~ 1,
    YEAR == 2013 & INCTOT < 36841 ~ 1,
    YEAR == 2014 & INCTOT < 38173 ~ 1,
    YEAR == 2015 & INCTOT < 39275 ~ 1,
    TRUE ~ 2
  )) 

df_samesex_inc_1 <- df_samesex_inc %>%
  filter(inc_class_year == 1) 


df_samesex_inc_2 <- df_samesex_inc %>%
  filter(inc_class_year == 2) 
```


##MARCEN BUT STRAIGHT COUPLES (ignore variable names, kept for matching purposes)
```{r, include = F}
gc()

df_samesex_st <- df_samesex %>%
  filter(in_samesex == 0) %>% 
  mutate(in_samesex_weighted = PERWT) %>% 
  filter(AGE >= 30 & AGE <= 64) %>%
  filter(YEAR >= 2001 & YEAR <= 2015) %>% 
  filter(MARST == 4 | MARST == 5 | MARST == 6) #includes those divorced, widowed, never married/single (NOTE: this excludes those married, separated, blank/missing)

#filter by sample
df_samesex_st_gen <- df_samesex_st #general sample
df_samesex_st_m <- df_samesex_st %>% #men only
                    filter(SEX.indv == 1)
df_samesex_st_f <- df_samesex_st %>% #women only
                    filter(SEX.indv == 2)
```


---W24 ARCHIVED--- final division and cleaning (see MM page 5 under the equation)

```{r, include = F}
# gc()
# 
# df <- df_samesex_mm_f
# 
# #make year totals excluding each state (no migrant restrictions)
# 
# df_yeartotals <- df
# 
# state_names <- unique(df$state_name)
# 
# indv_names <- paste(state_names, "excluded_indv", sep = "_")
# 
# for(i in 1:length(state_names)) {
#   df_yeartotals <- df_yeartotals %>% 
#     mutate(column = if_else(state_name != state_names[i], in_samesex_weighted, 0)) %>%
#     rename(!!indv_names[i] := column) 
# }
# 
# gc()
# total_names <- paste(state_names, "excluded_total", sep = "_")
# 
# for(i in 1:length(state_names)) {
#   col_name <- indv_names[i] 
#   
#   df_yeartotals <- df_yeartotals %>% 
#     group_by(YEAR) %>% 
#     mutate(columned = sum(!!sym(col_name), na.rm = T)) %>% 
#     ungroup() %>%
#     rename(!!total_names[i] := columned)
# }
# 
# df_yeartotals <- df_yeartotals %>%
#   select(YEAR, state_name, AL_excluded_total:WY_excluded_total)
# 
# #make state year totals
# 
# df_statetotals <- df %>%
#   filter(MIGPLAC1 > 0 & MIGPLAC1 < 57) %>%
#   group_by(YEAR, state_name) %>%
#   summarize(state_total = sum(in_samesex_weighted, na.rm = T)) %>%
#   ungroup()
# 
# #some states don't exist here so convert later to 0
# 
# #set up division
# 
# merged_totals <- df_statetotals %>%
#   left_join(df_yeartotals, join_by(YEAR, state_name), multiple = "any") #get rid of duplicates
# 
# samesex_staterate <- merged_totals %>% 
#   pivot_wider(names_from = state_name, values_from = state_total) %>%
#   mutate_all(~replace(., is.na(.), 0)) #coerce missing values to 0
# 
# #do the division
# 
# for(i in 1:length(state_names)) {
#   state = state_names[i]
#   total = total_names[i]
#   
#   samesex_staterate <- samesex_staterate %>% 
#     mutate(!!state := !!sym(state)/!!sym(total))
#   
# }
# 
# samesex_staterate <- samesex_staterate %>%
#   select(YEAR, AK:WY) %>%
#   pivot_longer(AK:WY, names_to = "state_name") %>%
#   rename(staterate_year = value) %>%
#   mutate(staterate_year = staterate_year * 100)
# rm(df, merged_totals, df_statetotals, df_yeartotals, col_name, i, indv_names, state, state_names, total, total_names)
# 
# gc()
# 
# 
# #ID key variables to merge back in
# extra_variables <- df_samesex %>% 
#   distinct(YEAR, state_name, time_since_legalization)
# 
# #merge back in 
# 
# output <-samesex_staterate %>%
#   left_join(extra_variables, join_by(YEAR, state_name)) %>%
#   left_join(FRED, join_by(YEAR, state_name))
# 
# #make dummy variables for 1-2, 3-4, 5-6, >7 years 
# output <- output %>%
#   mutate(time_leg_chunked = case_when(
#     time_since_legalization == 1 ~ 2,
#     time_since_legalization == 2 ~ 2,
#     time_since_legalization == 3 ~ 4,
#     time_since_legalization == 4 ~ 4,
#     time_since_legalization == 5 ~ 6,
#     time_since_legalization == 6 ~ 6,
#     time_since_legalization >= 7 ~ 7,
#   ))
# 
# 
# 
# rm(samesex_staterate, extra_variables)

#next steps: functionalize, adjust whether states are dropped by lack of same-sex individuals or lack of same-sex migrants, try building new dataframe technique (building row by row) instead of pivot technique
```

---W24 ARCHIVED--- Export
```{r, include = F}
# gc()
# write.csv(output, "C:\\Users\\njrich\\Desktop\\same-sex-migration\\inc_2.csv")
# write_dta(output, "C:\\Users\\njrich\\Desktop\\same-sex-migration\\inc_2.dta") #fill in rest by gender

```

--- F24 version (still expect differences from MM- see count differences remaining)--- final division and cleaning (see MM page 5 under the equation)
```{r, include = F}
# code
```

Summary statistics testing
```{r, include = F}
# code
```

Future regressions
```{r, include = F}
# code
```
