---
title: "figures"
output: html_document
date: "2024-12-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages (now many unnecessary)
```{r, include = F}
gc()
library(data.table) #for data loading
library(tidyverse) #for data work
install.packages("janitor")
library(janitor) #for clean_names
install.packages("srvyr")
library(srvyr) #for weighting
library(haven) #for reading in dta
install.packages("fastDummies")
library(fastDummies) #for region table
install.packages("kableExtra") #to make more complicated output tables for LATEX
library(kableExtra)

install.packages("jtools") #for ggplot visualization
library(jtools)

install.packages("ggpubr")
library(ggpubr) #to arrange plots together
```

Download clean data
```{r, include = F}
gc()
#windows
# clean_df <- read_dta("C:\\Users\\njrich\\Downloads\\clean_dataframe.dta") %>%
#   clean_names() %>%
#   mutate(across(where(is.labelled), ~ as.character(as_factor(.)))) #hopefully this doesn't break things

# MA_df <- read_dta("C:\\Users\\njrich\\Downloads\\MA_base_df.dta") %>%
#   clean_names() %>%
#   mutate(across(where(is.labelled), ~ as.character(as_factor(.)))) #hopefully this doesn't break things
#mac
clean_df <- read_dta("/Users/njrich/Downloads/clean_dataframe.dta") %>%
  clean_names() %>%
  mutate(across(where(is.labelled), ~ as.character(as_factor(.))))


#oop get rid of variables not using
```

Make survey object
```{r, include = F}
gc()

summary_stats <- as_survey(clean_df, weights = perwt) #note: exclude rep_wts so standard errors are inaccurate (file otherwise insanely big- talk someone about it)

#MA_summary_stats <- as_survey(MA_df, weights = perwt)
```

Summary stats - general population 
```{r, include = F}
gc()
#count those in/not in same sex
count_samesex <- summary_stats %>%
  group_by(year) %>%
  summarize(prop_samesex = survey_mean(in_samesex, na.rm = T)) %>%
  select(-prop_samesex_se)


gc()
prop_newlegal_samesex <- summary_stats %>% #dropped prior year; see sample changes each year make it hard to do this although accurate to my regression
  group_by(year, in_samesex) %>%
  summarize(prop_in_newlegal = survey_mean((1-expost_old_legal), na.rm = T)) %>%
  select(-prop_in_newlegal_se, -prop_from_newlegal_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "samesex", "oppositesex")) %>%
  pivot_wider(names_from = in_samesex, values_from = prop_in_newlegal) %>%
  ungroup

gc()
#create joint table
joint_overall_table <- count_male_female %>%
  left_join(prop_newlegal_samesex, by = "year") %>%
  mutate(across(-year, ~ . * 100)) %>%
  setNames(c("Year", "% Same-Sex", "Same-Sex", "Opposite-Sex")) %>%
  kable(caption = "Summary Statistics for Overall Population", format = "latex", booktabs = T, digits = 0) %>%
  add_header_above(c(" " = 2, "% in \textit{Obergefell}-legalized States" = 2))
#export joint table
writeLines(joint_overall_table, "/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/overall_table.tex") #NOTE: THIS IS FOR MAC, WILL NEED TO ADJUST FOR WINDOWS

#note frac v percent
#hmm text wrapping
```

Summary stats by Relationship Type By Variable
```{r, include = F}
gc()
#count frac. migrant:
prop_mig_samesex <- summary_stats %>%
  group_by(year, in_samesex) %>%
  summarize(prop_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-prop_migrate_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "Same-Sex Migrate", "Opposite-Sex Migrate")) %>%
  pivot_wider(names_from = in_samesex, values_from = prop_migrate) %>%
  select(year, "Same-Sex Migrate", "Opposite-Sex Migrate")


gc()
#count male/female among those in same-sex relationships
prop_fem_samesex <- summary_stats %>%
  mutate(sex = ifelse(sex == 9, NA, sex - 1)) %>%
  group_by(year, in_samesex) %>%
  summarize(prop_fem = survey_mean(sex, na.rm = T)) %>%
  select(-prop_fem_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "Same-Sex Sex", "Opposite-Sex Sex")) %>%
  pivot_wider(names_from = in_samesex, values_from = prop_fem) %>%
  select(year, "Same-Sex Sex", "Opposite-Sex Sex")

gc()
#fraction with children table
#need to import variables for this

gc()
#mean age by same-sex/opposite-sex weighted, by year
mean_age_samesex <- summary_stats %>%
  group_by(year, in_samesex) %>%
  summarize(mean_age = survey_mean(age, na.rm = T)) %>%
  select(-mean_age_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "Same-Sex Age", "Opposite-Sex Age")) %>%
  pivot_wider(names_from = in_samesex, values_from = mean_age) %>%
  select(year, "Same-Sex Age", "Opposite-Sex Age")

gc()
#fraction >=4 years college attainment by same-sex/opposite-sex weighted, by year
frac_college_samesex <- summary_stats %>%
  mutate(four_years_college = ifelse(educ >= 10, 1, 0)) %>%
  group_by(year, in_samesex) %>%
  summarize(frac_college = survey_mean(four_years_college, na.rm = T)) %>%
  select(-frac_college_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "Same-Sex College", "Opposite-Sex College")) %>%
  pivot_wider(names_from = in_samesex, values_from = frac_college) %>%
  select(year, "Same-Sex College", "Opposite-Sex College")

gc()
#fraction white by same-sex/opposite-sex weighted, by year
frac_white_samesex <- summary_stats %>%
  mutate(white = ifelse(race == 1, 1, 0)) %>%
  group_by(year, in_samesex) %>%
  summarize(frac_white = survey_mean(white, na.rm = T)) %>%
  select(-frac_white_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "Same-Sex White", "Opposite-Sex White")) %>%
  pivot_wider(names_from = in_samesex, values_from = frac_white)  %>%
  select(year, "Same-Sex White", "Opposite-Sex White")

gc()
#mean income by same-sex/opposite-sex weighted, by year
mean_income_samesex <- summary_stats %>%
  group_by(year, in_samesex) %>%
  summarize(mean_income = survey_mean(inctot, na.rm = T)) %>%
  select(-mean_income_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "Same-Sex Inc", "Opposite-Sex Inc")) %>%
  pivot_wider(names_from = in_samesex, values_from = mean_income) %>%
  select(year, "Same-Sex Inc", "Opposite-Sex Inc")

#frac have children by same-sex/opposite-sex weighted, by year
frac_child_samesex <- summary_stats %>%
  group_by(year, in_samesex) %>%
  summarize(frac_child = survey_mean(has_child, na.rm = T)) %>%
  select(-frac_child_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "Same-Sex Child", "Opposite-Sex Child")) %>%
  pivot_wider(names_from = in_samesex, values_from = frac_child) %>%
  select(year, "Same-Sex Child", "Opposite-Sex Child")


gc()
#create joint tables
joint_var_table_1 <- prop_mig_samesex %>%
  left_join(prop_fem_samesex, by = "year") %>%
  left_join(frac_child_samesex, by = "year") %>%
  ungroup %>%
  mutate(across(-year, ~ . * 100)) %>%
  setNames(c("Year", "Same-Sex", "Opposite-Sex", "Same-Sex", "Opposite-Sex", "Same-Sex", "Opposite-Sex")) %>%
  kable(caption = "Summary Statistics for Main Variables (1)", format = "latex", booktabs = T, digits = 0) %>%
  add_header_above(c(" " = 1, "% Migrate" = 2, "% Female" = 2, "% Has Child" = 2))  

joint_var_table_2 <- mean_age_samesex %>%
  left_join(frac_college_samesex, by = "year") %>%
  left_join(frac_white_samesex, by = "year") %>%
  left_join(mean_income_samesex, by = "year") %>%
  ungroup %>%
  mutate(across(-all_of(c("year", "Same-Sex Age", "Opposite-Sex Age", "Same-Sex Inc", "Opposite-Sex Inc")), ~ . * 100)) %>%
  setNames(c("Year", "Same-Sex", "Opposite-Sex", "Same-Sex", "Opposite-Sex", "Same-Sex", "Opposite-Sex", "Same-Sex", "Opposite-Sex")) %>% #needs work
  kable(caption = "Summary Statistics for Main Variables (2)", format = "latex", booktabs = T, digits = 0) %>%
  add_header_above(c(" " = 1, "Mean Age" = 2, "% College" = 2, "% White" = 2, "Mean Income" = 2))

#export joint tables
writeLines(joint_var_table_1, "/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/var_table_1.tex") #for mac
writeLines(joint_var_table_2, "/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/var_table_2.tex") 

#watch column arranging
```

State Type Stats
```{r, include = F}
# gc()
# prop_newlegal_samesex <- summary_stats %>%
#   group_by(year, in_samesex) %>%
#   summarize(prop_in_newlegal = survey_mean((1-expost_old_legal), na.rm = T)) %>%
#   select(-prop_in_newlegal_se, -prop_from_newlegal_se) %>%
#   mutate(in_samesex = ifelse(in_samesex == 1, "samesex", "oppositesex")) %>%
#   pivot_wider(names_from = in_samesex, values_from = prop_in_newlegal) %>%
#   ungroup
# 
# prop_newlegal_samesex_export <- prop_newlegal_samesex %>%
#   mutate(across(-year, ~ . * 100)) %>%
#   setNames(c("Year", "Same-Sex", "Opposite-Sex", "Same-Sex", "Opposite-Sex")) %>% #needs work
#   kable(caption = "Sample by State Type", format = "latex", booktabs = T, digits = 0) %>%
#   add_header_above(c(" " = 1, "% To States Without State-Level Marriage Equality" = 2, "% From States Without State-Level Marriage Equality" = 2))
# 
# #export
# writeLines(prop_newlegal_samesex_export, "/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/prop_newlegal_samesex.tex") #for mac

#hmmm.... add SE?
```


Region Stats
```{r, include = F}
gc()
prop_region_samesex <- clean_df %>%
  mutate(region = expost_region) %>%
  dummy_cols(select_columns = "region", remove_first_dummy = F) %>%
  clean_names %>%
  as_survey(weights = perwt) %>%
  group_by(year, in_samesex) %>%
  summarize(prop_midwest = survey_mean(region_midwest, na.rm = T),
            prop_south = survey_mean(region_south, na.rm = T),
            prop_west = survey_mean(region_west, na.rm = T),
            prop_northeast = survey_mean(region_northeast, na.rm = T)) %>%
  select(-prop_midwest_se, -prop_south_se, -prop_west_se, -prop_northeast_se) %>%
  mutate(in_samesex = ifelse(in_samesex == 1, "samesex", "oppositesex")) %>%
  pivot_wider(names_from = in_samesex, values_from = prop_midwest:prop_northeast) %>%
  ungroup

regions_1 <- prop_region_samesex %>%
  select(year, prop_midwest_samesex, prop_midwest_oppositesex, prop_south_samesex, prop_south_oppositesex) %>%
  mutate(across(-year, ~ . * 100)) %>%
  setNames(c("Year", "Same-Sex", "Opposite-Sex", "Same-Sex", "Opposite-Sex")) %>% #needs work
  kable(caption = "Summary Statistics by Region (1)", format = "latex", booktabs = T, digits = 0) %>%
  add_header_above(c(" " = 1, "% in the Midwest" = 2, "% in the South" = 2))

regions_2 <- prop_region_samesex %>%
  select(year, prop_west_samesex, prop_west_oppositesex, prop_northeast_samesex, prop_northeast_oppositesex) %>%
  mutate(across(-year, ~ . * 100)) %>%
  setNames(c("Year", "Same-Sex", "Opposite-Sex", "Same-Sex", "Opposite-Sex")) %>% #needs work
  kable(caption = "Summary Statistics by Region (2)", format = "latex", booktabs = T, digits = 0) %>%
  add_header_above(c(" " = 1, "% in the West" = 2, "% in the Northeast" = 2))

#export joint tables
writeLines(regions_1, "/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/regions_1.tex") #for mac
writeLines(regions_2, "/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/regions_2.tex") 
```

Main Model Trends/Differences Charts
```{r, include = F}
gc()
#Pull Factor
#df
post_migrant_trends_df <- summary_stats %>%
  mutate(subgroup = case_when(
    in_samesex == 1 & expost_old_legal == 1 ~ "samesex_oldlegal",
    in_samesex == 0 & expost_old_legal == 1 ~ "oppositesex_oldlegal",
    in_samesex == 1 & expost_old_legal == 0 ~ "samesex_newlegal",
    in_samesex == 0 & expost_old_legal == 0 ~ "oppositesex_newlegal"
  )) %>%
  group_by(year, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) 

#table for export
post_migrant_trendsdiffs_table <- post_migrant_trends_df %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  select(year, samesex_oldlegal, samesex_newlegal, samesex_diff, oppositesex_oldlegal, oppositesex_newlegal, oppositesex_diff)


#trends chart for export
# gc()
# post_migrant_trends_chart <- post_migrant_trends_df %>%
#   ggplot() +
#     geom_line(aes(x = year, y = frac_migrate, linetype = subgroup)) + 
#     labs(title = "Pull Factor Trends Visualized",
#          x = "Year",
#          y = "Fraction Migrated",
#          linetype = "Subgroup") + #maybe move title to plot itself
#     geom_vline(xintercept = 2015, color = "black", linetype = "dashed") + #bc factor issues
#     scale_linetype_manual(values = c("samesex_oldlegal" = "solid", "oppositesex_oldlegal" = "dotted", "samesex_newlegal" = "longdash", "oppositesex_newlegal" = "dotdash"),
#                      labels = c("samesex_oldlegal" = "Same-Sex Individuals in Locally Legalized States", "oppositesex_oldlegal" = "Opposite-Sex Individuals in Locally Legalized States", "samesex_newlegal" = "Same-Sex Individuals in Federally Legalized States", "oppositesex_newlegal" = "Opposite-Sex Individuals in Federally Legalized States")) +
#   scale_x_continuous(breaks = 2011:2019) +
#   theme_apa() #can always go back to theme
# 
# post_migrant_trends_chart

#diffs chart for export
gc()
post_migrant_diffs_chart <- post_migrant_trendsdiffs_table %>%
  select(year, samesex_diff, oppositesex_diff) %>%
  pivot_longer(cols = c(samesex_diff, oppositesex_diff), names_to = "subgroup", values_to= "diffs") %>%
  mutate(diffs = -diffs) %>% #modified in light of model changes
  ggplot() +
    geom_line(aes(x = year, y = diffs, linetype = subgroup)) + 
    labs(title = "Pull Factor Triple Regression Visualized",
         x = "Year",
         y = "Relative Migration to Obergefell States",
         linetype = "Subgroup") + #maybe move title to plot itself
    geom_vline(xintercept = 2015, color = "black", linetype = "dotted") + #bc factor issues
    scale_linetype_manual(values = c("samesex_diff" = "solid", "oppositesex_diff" = "longdash"),
                     labels = c("samesex_diff" = "Same-Sex Individuals", "oppositesex_diff" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa() #can always go back to theme

post_migrant_diffs_chart

#ante
#df
ante_migrant_trends_df <- summary_stats %>%
  mutate(subgroup = case_when(
    in_samesex == 1 & exante_old_legal == 1 ~ "samesex_oldlegal",
    in_samesex == 0 & exante_old_legal == 1 ~ "oppositesex_oldlegal",
    in_samesex == 1 & exante_old_legal == 0 ~ "samesex_newlegal",
    in_samesex == 0 & exante_old_legal == 0 ~ "oppositesex_newlegal"
  )) %>%
  group_by(year, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) 

#table for export
ante_migrant_trendsdiffs_table <- ante_migrant_trends_df %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  select(year, samesex_oldlegal, samesex_newlegal, samesex_diff, oppositesex_oldlegal, oppositesex_newlegal, oppositesex_diff)

#trends chart for export
# gc()
# ante_migrant_trends_chart <- ante_migrant_trends_df %>%
#   ggplot() +
#     geom_line(aes(x = year, y = frac_migrate, linetype = subgroup)) +
#   labs(title = "Push Factor Trends Visualized",
#         x = "Year",
#         y = "Fraction Migrated",
#         linetype = "Subgroup") +
#   geom_vline(xintercept = 2015, color = "black", linetype = "dashed") + #bc factor issues
#   scale_linetype_manual(values = c("samesex_oldlegal" = "solid", "oppositesex_oldlegal" = "dotted", "samesex_newlegal" = "longdash", "oppositesex_newlegal" = "dotdash"),
#                      labels = c("samesex_oldlegal" = "Same-Sex Individuals in Locally Legalized States", "oppositesex_oldlegal" = "Opposite-Sex Individuals in Locally Legalized States", "samesex_newlegal" = "Same-Sex Individuals in Federally Legalized States", "oppositesex_newlegal" = "Opposite-Sex Individuals in Federally Legalized States")) +
#   scale_x_continuous(breaks = 2011:2019) +
#   theme_apa()
# 
# ante_migrant_trends_chart

#diffs chart for export
gc()
ante_migrant_diffs_chart <- ante_migrant_trendsdiffs_table %>%
  select(year, samesex_diff, oppositesex_diff) %>%
  pivot_longer(cols = c(samesex_diff, oppositesex_diff), names_to = "subgroup", values_to= "diffs") %>%
  mutate(diffs = -diffs) %>% #modified in light of model changes
  ggplot() +
    geom_line(aes(x = year, y = diffs, linetype = subgroup)) + 
    labs(title = "Push Factor Triple Regression Visualized",
         x = "Year",
         y = "Relative Migration from Obergefell States",
         linetype = "Subgroup") + #maybe move title to plot itself
    geom_vline(xintercept = 2015, color = "black", linetype = "dashed") + #bc factor issues
    scale_linetype_manual(values = c("samesex_diff" = "solid", "oppositesex_diff" = "dashed"),
                     labels = c("samesex_diff" = "Same-Sex Individuals", "oppositesex_diff" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa() #can always go back to theme

ante_migrant_diffs_chart


#export charts
#ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/post_trends.png", plot = post_migrant_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/post_diffs.png", plot = post_migrant_diffs_chart)
#ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/ante_trends.png", plot = ante_migrant_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/ante_diffs.png", plot = ante_migrant_diffs_chart)


#export joint table
# joint_trendsdiffs_table <- post_migrant_trendsdiffs_table %>%
#   left_join(ante_migrant_trendsdiffs_table, by = "year") %>% 
#   setNames(c("Year", "Locally Legalized", "Federally Legalized", "Difference", "Locally Legalized", "Federally Legalized", "Difference", "Locally Legalized", "Federally Legalized", "Difference", "Locally Legalized", "Federally Legalized", "Difference")) %>%
#   kable(caption = "Migration Trends Data", format = "latex", booktabs = T, digits = 0) %>%
#   add_header_above(c(" " = 1, "Same-Sex" = 2, " " = 1, "Opposite-Sex" = 2, " " = 1, "Same-Sex" = 2, " " = 1, "Opposite-Sex" = 2, " " = 1))  %>% #hmm reorganize to make more sense?
#   add_header_above(c(" " = 1, "Fraction Moving To" = 6, "Fraction Moving From" = 6))
# 
# writeLines(joint_trendsdiffs_table, "/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/trendsdiffs_table.tex")
gc()
#oop functionalize
#watch issues with code, better labeling and organization
#separate these boxes?
#discrepancies in plots to test

#for charts to look good, just take screenshots at the moment
```



Flow Heterogeneity Charts
```{r, include = F}
gc()
#make dfs
post_migrant_trends_df <- summary_stats %>%
  mutate(subgroup = case_when(
    in_samesex == 1 & expost_old_legal == 1 ~ "samesex_oldlegal",
    in_samesex == 0 & expost_old_legal == 1 ~ "oppositesex_oldlegal",
    in_samesex == 1 & expost_old_legal == 0 ~ "samesex_newlegal",
    in_samesex == 0 & expost_old_legal == 0 ~ "oppositesex_newlegal"
  ))

ante_migrant_trends_df <- summary_stats %>%
  mutate(subgroup = case_when(
    in_samesex == 1 & exante_old_legal == 1 ~ "samesex_oldlegal",
    in_samesex == 0 & exante_old_legal == 1 ~ "oppositesex_oldlegal",
    in_samesex == 1 & exante_old_legal == 0 ~ "samesex_newlegal",
    in_samesex == 0 & exante_old_legal == 0 ~ "oppositesex_newlegal"
  )) 


#make functions to make charts

make_flowtrend_chart <- function(df, sel_title, frac_migrate) {
  chart <- df %>%
  ggplot() +
    geom_line(aes(x = year, y = {{ frac_migrate }}, linetype = subgroup)) + 
    labs(title = sel_title,
         x = "Year",
         y = "Fraction Migrated",
         linetype = "Subgroup") +
    geom_vline(xintercept = 2015, color = "black", linetype = "dashed") + #bc factor issues
    scale_linetype_manual(values = c("samesex_oldlegal" = "solid", "oppositesex_oldlegal" = "dotted", "samesex_newlegal" = "longdash", "oppositesex_newlegal" = "dotdash"),
                     labels = c("samesex_oldlegal" = "Same-Sex Individuals in Locally Legalized States", "oppositesex_oldlegal" = "Opposite-Sex Individuals in Locally Legalized States", "samesex_newlegal" = "Same-Sex Individuals in Federally Legalized States", "oppositesex_newlegal" = "Opposite-Sex Individuals in Federally Legalized States")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa()
  
  return(chart)
}

make_flowdiffs_chart_post <- function(df, sel_title) {
  chart <- df %>%
  select(year, samesex_diff, oppositesex_diff) %>%
  pivot_longer(cols = c(samesex_diff, oppositesex_diff), names_to = "subgroup", values_to= "diffs") %>%
    mutate(diffs = -diffs) %>% #adjusted due to model changes
  ggplot() +
    geom_line(aes(x = year, y = diffs, linetype = subgroup)) + 
    labs(title = sel_title,
         x = "Year",
         y = "Relative Migration to Obergefell States", #ambiguous for purposes of from v. to
         linetype = "Subgroup") + 
    geom_vline(xintercept = 2015, color = "black", linetype = "dotted") + 
    scale_linetype_manual(values = c("samesex_diff" = "solid", "oppositesex_diff" = "longdash"),
                     labels = c("samesex_diff" = "Same-Sex Individuals", "oppositesex_diff" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa()
  return(chart)
}

make_flowdiffs_chart_ante <- function(df, sel_title) {
  chart <- df %>%
  select(year, samesex_diff, oppositesex_diff) %>%
  pivot_longer(cols = c(samesex_diff, oppositesex_diff), names_to = "subgroup", values_to= "diffs") %>%
    mutate(diffs = -diffs) %>% #adjusted due to model changes
  ggplot() +
    geom_line(aes(x = year, y = diffs, linetype = subgroup)) + 
    labs(title = sel_title,
         x = "Year",
         y = "Relative Migration from Obergefell States", #ambiguous for purposes of from v. to
         linetype = "Subgroup") + 
    geom_vline(xintercept = 2015, color = "black", linetype = "dotted") + 
    scale_linetype_manual(values = c("samesex_diff" = "solid", "oppositesex_diff" = "longdash"),
                     labels = c("samesex_diff" = "Same-Sex Individuals", "oppositesex_diff" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa()
  return(chart)
}

#Post
#From Fed
fed_post_df <- post_migrant_trends_df %>% #should help code run faster
  group_by(year, subgroup) %>%
  summarize(frac_ffed_migrate = survey_mean(fromfed_migrant, na.rm = T)) %>%
  select(-frac_ffed_migrate_se) %>%
  ungroup
  
#fed_post_trends_chart <- fed_post_df %>%
#  make_flowtrend_chart(sel_title = "Pull Factor Trends From Not/Fed Legalized", frac_migrate = frac_ffed_migrate) 

fed_post_diffs_chart <- fed_post_df %>%
  pivot_wider(names_from = subgroup, values_from = frac_ffed_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  make_flowdiffs_chart_post(sel_title = "Pull Factor Diffs From Obergefell States") 
fed_post_diffs_chart
gc()
#From Local (of interest)
local_post_df <- post_migrant_trends_df %>% #should help code run faster
  group_by(year, subgroup) %>%
  summarize(frac_flocal_migrate = survey_mean(fromlocal_migrant, na.rm = T)) %>%
  select(-frac_flocal_migrate_se) %>%
  ungroup
  
#local_post_trends_chart <- local_post_df %>%
#  make_flowtrend_chart(sel_title = "Pull Factor Trends From Locally Legalized", frac_migrate = frac_flocal_migrate)

local_post_diffs_chart <- local_post_df %>%
  pivot_wider(names_from = subgroup, values_from = frac_flocal_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  make_flowdiffs_chart_post(sel_title = "Pull Factor Diffs from non-Obergefell States")
local_post_diffs_chart
gc()
#Post Stay (note: should technically be the same Pull Factor/Push Factor frame except for the fact states can change legal state between years)
stay_post_df <- post_migrant_trends_df %>% #note: should really only be one value
  group_by(year, subgroup) %>%
  summarize(frac_fstay = survey_mean(1 - migrant, na.rm = T)) %>%
  select(-frac_fstay_se) %>%
  ungroup
  
#stay_post_trends_chart <- stay_post_df %>%
#  make_flowtrend_chart(sel_title = "Pull Factor Trends From Stay", frac_migrate = frac_fstay)

stay_post_diffs_chart <- stay_post_df %>%
  pivot_wider(names_from = subgroup, values_from = frac_fstay) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  make_flowdiffs_chart_post(sel_title = "Pull Factor Diffs from the Same State")
stay_post_diffs_chart
#Ante
gc()
#To Fed
fed_ante_df <- ante_migrant_trends_df %>% #should help code run faster
  group_by(year, subgroup) %>%
  summarize(frac_tfed_migrate = survey_mean(tofed_migrant, na.rm = T)) %>%
  select(-frac_tfed_migrate_se) %>%
  ungroup
  
#fed_ante_trends_chart <- fed_ante_df %>%
#  make_flowtrend_chart(sel_title = "Push Factor Trends To Not/Fed Legalized", frac_migrate = frac_tfed_migrate) 

fed_ante_diffs_chart <- fed_ante_df %>%
  pivot_wider(names_from = subgroup, values_from = frac_tfed_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  make_flowdiffs_chart_ante(sel_title = "Push Factor Diffs to Obergefell States") 
fed_ante_diffs_chart
gc()
#To Local (of interest)
local_ante_df <- ante_migrant_trends_df %>% #should help code run faster
  group_by(year, subgroup) %>%
  summarize(frac_tlocal_migrate = survey_mean(tolocal_migrant, na.rm = T)) %>%
  select(-frac_tlocal_migrate_se) %>%
  ungroup
  
#local_ante_trends_chart <- local_ante_df %>%
#  make_flowtrend_chart(sel_title = "Push Factor Trends To Locally Legalized", frac_migrate = frac_tlocal_migrate)

local_ante_diffs_chart <- local_ante_df %>%
  pivot_wider(names_from = subgroup, values_from = frac_tlocal_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  make_flowdiffs_chart_ante(sel_title = "Push Factor Diffs to non-Obergefell States")
local_ante_diffs_chart
#Ante Stay (note: should technically be the same Pull Factor/Push Factor frame except for the fact states can change legal state between years)
gc()
stay_ante_df <- ante_migrant_trends_df %>% #note: should really only be one value
  group_by(year, subgroup) %>%
  summarize(frac_tstay = survey_mean(1 - migrant, na.rm = T)) %>%
  select(-frac_tstay_se) %>%
  ungroup
  
#stay_ante_trends_chart <- stay_ante_df %>%
#  make_flowtrend_chart(sel_title = "Push Factor Trends To Stay", frac_migrate = frac_tstay)

stay_ante_diffs_chart <- stay_ante_df %>%
  pivot_wider(names_from = subgroup, values_from = frac_tstay) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  make_flowdiffs_chart_ante(sel_title = "Push Factor Diffs to the Same State")
stay_ante_diffs_chart

#Combine
#flows_post_trends_chart <- ggarrange(fed_post_trends_chart, local_post_trends_chart, stay_post_trends_chart, nrow = 1, common.legend = T, legend = "right")
#flows_ante_trends_chart <- ggarrange(fed_ante_trends_chart, local_ante_trends_chart, stay_ante_trends_chart, nrow = 1, common.legend = T, legend = "right")
flows_post_diffs_chart_app <- ggarrange(fed_post_diffs_chart, stay_post_diffs_chart, nrow = 2, common.legend = T, legend = "right")
flows_ante_diffs_chart_app <- ggarrange(fed_ante_diffs_chart, stay_ante_diffs_chart, nrow = 2, common.legend = T, legend = "right")

#Save and Export
#ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/flows_post_trends.png", plot = flows_post_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/flows_post_diffs_app.png", plot = flows_post_diffs_chart_app)
#ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/flows_ante_trends.png", plot = flows_ante_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/flows_ante_diffs_app.png", plot = flows_ante_diffs_chart_app)
#oop watch cohesive export size need to unify
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/flows_post_diffs.png", plot = local_post_diffs_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/flows_ante_diffs.png", plot = local_ante_diffs_chart)


#more functionalize?
#watch naming
#darn naming conventions
#watch typos
```


Main heterogeneity charts (first output: trend, second output: diff; can adjust first step chart making to make code take half the time)
```{r, include = F}
gc()
#make dfs
post_migrant_trends_df <- summary_stats %>%
  mutate(subgroup = case_when(
    in_samesex == 1 & expost_old_legal == 1 ~ "samesex_oldlegal",
    in_samesex == 0 & expost_old_legal == 1 ~ "oppositesex_oldlegal",
    in_samesex == 1 & expost_old_legal == 0 ~ "samesex_newlegal",
    in_samesex == 0 & expost_old_legal == 0 ~ "oppositesex_newlegal"
  ))

ante_migrant_trends_df <- summary_stats %>%
  mutate(subgroup = case_when(
    in_samesex == 1 & exante_old_legal == 1 ~ "samesex_oldlegal",
    in_samesex == 0 & exante_old_legal == 1 ~ "oppositesex_oldlegal",
    in_samesex == 1 & exante_old_legal == 0 ~ "samesex_newlegal",
    in_samesex == 0 & exante_old_legal == 0 ~ "oppositesex_newlegal"
  )) 

#make functions to make charts

make_trend_chart <- function(df, sel_title, facet, facet_titles) {
  chart <- df %>%
  ggplot() +
    geom_line(aes(x = year, y = frac_migrate, linetype = subgroup)) + 
    labs(title = sel_title,
         x = "Year",
         y = "Fraction Migrated",
         linetype = "Subgroup") +
    geom_vline(xintercept = 2015, color = "black", linetype = "dashed") + #bc factor issues
    scale_linetype_manual(values = c("samesex_oldlegal" = "solid", "oppositesex_oldlegal" = "dotted", "samesex_newlegal" = "longdash", "oppositesex_newlegal" = "dotdash"),
                     labels = c("samesex_oldlegal" = "Same-Sex Individuals in Locally Legalized States", "oppositesex_oldlegal" = "Opposite-Sex Individuals in Locally Legalized States", "samesex_newlegal" = "Same-Sex Individuals in Federally Legalized States", "oppositesex_newlegal" = "Opposite-Sex Individuals in Federally Legalized States")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa() +
  facet_wrap(reformulate(facet), labeller = as_labeller(facet_titles))
  
  return(chart)
}

facet_make_diffs_chart_post <- function(df, sel_title, group, facet, facet_titles) {
  chart <- df %>%
  select(year, {{ group }}, samesex_diff, oppositesex_diff) %>%
  pivot_longer(cols = c(samesex_diff, oppositesex_diff), names_to = "subgroup", values_to= "diffs") %>%
  mutate(diffs = -diffs) %>% #adjusted due to model changes
  ggplot() +
    geom_line(aes(x = year, y = diffs, linetype = subgroup)) + 
    labs(title = sel_title,
         x = "Year",
         y = "Relative Migration to Obergefell States", #ambiguous for purposes of from v. to
         linetype = "Subgroup") + 
    geom_vline(xintercept = 2015, color = "black", linetype = "dotted") + 
    scale_linetype_manual(values = c("samesex_diff" = "solid", "oppositesex_diff" = "longdash"),
                     labels = c("samesex_diff" = "Same-Sex Individuals", "oppositesex_diff" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa() +
  facet_wrap(reformulate(facet), labeller = as_labeller(facet_titles), nrow = 2) #added nrow term
  return(chart)
}

facet_make_diffs_chart_ante <- function(df, sel_title, group, facet, facet_titles) {
  chart <- df %>%
  select(year, {{ group }}, samesex_diff, oppositesex_diff) %>%
  pivot_longer(cols = c(samesex_diff, oppositesex_diff), names_to = "subgroup", values_to= "diffs") %>%
  mutate(diffs = -diffs) %>% #adjusted due to model changes
  ggplot() +
    geom_line(aes(x = year, y = diffs, linetype = subgroup)) + 
    labs(title = sel_title,
         x = "Year",
         y = "Relative Migration from Obergefell States", #ambiguous for purposes of from v. to
         linetype = "Subgroup") + 
    geom_vline(xintercept = 2015, color = "black", linetype = "dotted") + 
    scale_linetype_manual(values = c("samesex_diff" = "solid", "oppositesex_diff" = "longdash"),
                     labels = c("samesex_diff" = "Same-Sex Individuals", "oppositesex_diff" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa() +
  facet_wrap(reformulate(facet), labeller = as_labeller(facet_titles), nrow = 2) #added nrow term
  return(chart)
}
#see functions saved earlier - flowdiffs ones which should work if need non-facet tables

gc()
###Split by sex
#Post
# sex_post_trends_chart <- post_migrant_trends_df %>%
#   group_by(year, sex, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   mutate(sex_label = if_else(sex == 1, "male", "female")) %>%
#   make_trend_chart(sel_title = "Pull Factor Trends by Gender", facet = "sex_label", facet_titles = c("male" = "Men", "female" = "Women")) #use male/female?

sex_post_diffs_df <- post_migrant_trends_df %>%
  group_by(year, sex, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  mutate(sex_label = if_else(sex == 1, "male", "female")) 

men_post_diffs_chart <- sex_post_diffs_df %>%
  filter(sex_label == "male") %>%
  make_flowdiffs_chart_post(sel_title = "Pull Factor Diffs for Men") #use male/female?
men_post_diffs_chart
women_post_diffs_chart <- sex_post_diffs_df %>%
  filter(sex_label == "female") %>%
  make_flowdiffs_chart_post(sel_title = "Pull Factor Diffs for Women") #use male/female?
women_post_diffs_chart

gc()
#Ante
# sex_ante_trends_chart <- ante_migrant_trends_df %>%
#   group_by(year, sex, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   mutate(sex_label = if_else(sex == 1, "male", "female")) %>%
#   make_trend_chart(sel_title = "Push Factor Trends by Gender", facet = "sex_label", facet_titles = c("male" = "Men", "female" = "Women")) #use male/female?

sex_ante_diffs_df <- post_migrant_trends_df %>%
  group_by(year, sex, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  mutate(sex_label = if_else(sex == 1, "male", "female"))

men_ante_diffs_chart <- sex_ante_diffs_df %>%
  filter(sex_label == "male") %>%
  make_flowdiffs_chart_ante(sel_title = "Push Factor Diffs for Men") #use male/female?
men_ante_diffs_chart
women_ante_diffs_chart <- sex_ante_diffs_df %>%
  filter(sex_label == "female") %>%
  make_flowdiffs_chart_ante(sel_title = "Push Factor Diffs for Women") #use male/female?
women_ante_diffs_chart
#takeaways: WOMEN ARE DRIVING Pull Factor/Push Factor DIFFERENCES YES MAKES SENSE OH LORDY

gc()
###Split by Age
#Post
# age_post_trends_chart <- post_migrant_trends_df %>%
#   mutate(age_groups = case_when(
#     age < 23 ~ "<23",
#     age >= 23 & age < 33 ~ "23-32",
#     age >= 33 & age < 63 ~ "33-62",
#     age >= 63 ~ ">62")) %>% #make these groups better
#   group_by(year, age_groups, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Pull Factor Trends by Age", facet = "age_groups", facet_titles = c("<23" = "<23", "23-32" = "23-32", "33-62" = "33-62", ">62" = ">62"))

age_post_diffs_chart <- post_migrant_trends_df %>%
  mutate(age_groups = case_when(
    age < 23 ~ "<23",
    age >= 23 & age < 33 ~ "23-32",
    age >= 33 & age < 63 ~ "33-62",
    age >= 63 ~ ">62")) %>% #make these groups better
  group_by(year, age_groups, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  facet_make_diffs_chart_post(sel_title = "Pull Factor Diffs by Age", group = age_groups, facet = "age_groups", c("<23" = "<23", "23-32" = "23-32", "33-62" = "33-62", ">62" = ">62"))
age_post_diffs_chart

gc()
#Ante
# age_ante_trends_chart <- ante_migrant_trends_df %>%
#   mutate(age_groups = case_when(
#     age < 23 ~ "<23",
#     age >= 23 & age < 33 ~ "23-32",
#     age >= 33 & age < 63 ~ "33-62",
#     age >= 63 ~ ">62")) %>% #make these groups better
#   group_by(year, age_groups, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Push Factor Trends by Age", facet = "age_groups", facet_titles = c("<23" = "<23", "23-32" = "23-32", "33-62" = "33-62", ">62" = ">62"))

age_ante_diffs_chart <- ante_migrant_trends_df %>%
  mutate(age_groups = case_when(
    age < 23 ~ "<23",
    age >= 23 & age < 33 ~ "23-32",
    age >= 33 & age < 63 ~ "33-62",
    age >= 63 ~ ">62")) %>% #make these groups better
  group_by(year, age_groups, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  facet_make_diffs_chart_ante(sel_title = "Push Factor Diffs by Age", group = age_groups, facet = "age_groups", c("<23" = "<23", "23-32" = "23-32", "33-62" = "33-62", ">62" = ">62"))
age_ante_diffs_chart
#takeaways: minimal meaning, area most of interest is 23-32 (most mobile, period most likely to marry)

###Split by Educ
gc()
#Post
# educ_post_trends_chart <- post_migrant_trends_df %>%
#   mutate(college = ifelse(educ >= 10, "college", "notcollege")) %>%
#   group_by(year, college, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Pull Factor Trends by Education", facet = "college", facet_titles = c("college" = ">=4 Years College", "notcollege" = "<4 Years College"))

educ_post_diffs_chart <- post_migrant_trends_df %>%
  mutate(college = ifelse(educ >= 10, "college", "notcollege")) %>%
  group_by(year, college, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  facet_make_diffs_chart_post(sel_title = "Pull Factor Diffs by Education", group = college, facet = "college", facet_titles = c("college" = ">=4 Years College", "notcollege" = "<4 Years College"))
educ_post_diffs_chart
gc()
#Ante
# educ_ante_trends_chart <- ante_migrant_trends_df %>%
#   mutate(college = ifelse(educ >= 10, "college", "notcollege")) %>%
#   group_by(year, college, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Push Factor Trends by Education", facet = "college", facet_titles = c("college" = ">=4 Years College", "notcollege" = "<4 Years College"))

educ_ante_diffs_chart <- post_migrant_trends_df %>%
  mutate(college = ifelse(educ >= 10, "college", "notcollege")) %>%
  group_by(year, college, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  facet_make_diffs_chart_ante(sel_title = "Push Factor Diffs by Education", group = college, facet = "college", facet_titles = c("college" = ">=4 Years College", "notcollege" = "<4 Years College"))
educ_ante_diffs_chart
#takeaways: kind of noisy, doesn't seem to follow predicted (post) trend

###Split by Children
gc()
#Post
# child_post_trends_chart <- post_migrant_trends_df %>%
#   mutate(child = ifelse(has_child ==1, "yes", "no")) %>%
#   group_by(year, child, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Pull Factor Trends by Presence of Children", facet = "child", facet_titles = c("yes" = "Children Present", "no" = "Children not Present"))

child_post_diffs_chart <- post_migrant_trends_df %>%
  mutate(child = ifelse(has_child ==1, "yes", "no")) %>%
  group_by(year, child, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  facet_make_diffs_chart_post(sel_title = "Pull Factor Diffs by Presence of Children", group = child, facet = "child", facet_titles = c("yes" = "Children Present", "no" = "Children not Present"))
child_post_diffs_chart
gc()
#Ante
# child_ante_trends_chart <- ante_migrant_trends_df %>%
#   mutate(child = ifelse(has_child ==1, "yes", "no")) %>%
#   group_by(year, child, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Push Factor Trends by Presence of Children", facet = "child", facet_titles = c("yes" = "Children Present", "no" = "Children not Present"))

child_ante_diffs_chart <- post_migrant_trends_df %>%
  mutate(child = ifelse(has_child ==1, "yes", "no")) %>%
  group_by(year, child, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  facet_make_diffs_chart_ante(sel_title = "Push Factor Diffs by Presence of Children", group = child, facet = "child", facet_titles = c("yes" = "Children Present", "no" = "Children not Present"))
child_ante_diffs_chart
#takeaways: noisy; no child post chart kind of interesting/follows predictions (ante looks similar)

###Split by Income
gc()
#Post
# inc_post_trends_chart <- post_migrant_trends_df %>%
#   group_by(year) %>%
#   mutate(mean_income = sum(perwt * inctot, na.rm = T) / sum(perwt, na.rm = T)) %>% #note: applies to whole year population, not separated, watch mean v median, own bootleg weight calculations
#   ungroup() %>%
#   mutate(income_group = ifelse(inctot < mean_income, "below", "above")) %>%
#   group_by(year, income_group, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Pull Factor Trends by Income", facet = "income_group", facet_titles = c("above" = "Above Mean", "below" = "Below Mean"))

inc_post_diffs_chart <- post_migrant_trends_df %>%
    group_by(year) %>%
  mutate(mean_income = sum(perwt * inctot, na.rm = T) / sum(perwt, na.rm = T)) %>% #note: applies to whole year population, not separated, watch mean v median, own bootleg weight calculations
  ungroup() %>%
  mutate(income_group = ifelse(inctot < mean_income, "below", "above")) %>%
  group_by(year, income_group, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  facet_make_diffs_chart_post(sel_title = "Pull Factor Diffs by Income", group = income_group, facet = "income_group", facet_titles = c("above" = "Above Mean", "below" = "Below Mean"))
inc_post_diffs_chart

gc()
#Ante
# inc_ante_trends_chart <- ante_migrant_trends_df %>%
#     group_by(year) %>%
#   mutate(mean_income = sum(perwt * inctot, na.rm = T) / sum(perwt, na.rm = T)) %>% #note: applies to whole year population, not separated, watch mean v median, own bootleg weight calculations
#   ungroup() %>%
#   mutate(income_group = ifelse(inctot < mean_income, "below", "above")) %>%
#   group_by(year, income_group, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Push Factor Trends by Income", facet = "income_group", facet_titles = c("above" = "Above Mean", "below" = "Below Mean"))

inc_ante_diffs_chart <- post_migrant_trends_df %>%
    group_by(year) %>%
  mutate(mean_income = sum(perwt * inctot, na.rm = T) / sum(perwt, na.rm = T)) %>% #note: applies to whole year population, not separated, watch mean v median, own bootleg weight calculations
  ungroup() %>%
  mutate(income_group = ifelse(inctot < mean_income, "below", "above")) %>%
  group_by(year, income_group, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) %>%
  facet_make_diffs_chart_ante(sel_title = "Push Factor Diffs by Income", group = income_group, facet = "income_group", facet_titles = c("above" = "Above Mean", "below" = "Below Mean"))
inc_ante_diffs_chart
#takeaways: noisy, (post) kind of follows hypothesis

gc()
###Split by Region
#Post
# gc()
# region_post_trends_chart <- post_migrant_trends_df %>%
#   group_by(year, expost_region, subgroup) %>% #be careful
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Pull Factor Trends by Region", facet = "expost_region", facet_titles = c("Midwest" = "Midwest", "West" = "West", "South" = "South", "Northeast" = "Northeast"))

gc()
region_post_diffs_df <- post_migrant_trends_df %>%
  group_by(year, expost_region, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal)

midwest_south_post_diffs_chart <- region_post_diffs_df %>%
  filter(expost_region %in% c("Midwest", "South")) %>%
  facet_make_diffs_chart_post(sel_title = "Pull Factor Diffs in the Midwest and South", group = expost_region, facet = "expost_region", facet_titles = c("Midwest" = "Midwest", "South" = "South"))
midwest_south_post_diffs_chart

northeast_west_post_diffs_chart <- region_post_diffs_df %>%
  filter(expost_region %in% c("Northeast", "West")) %>%
  facet_make_diffs_chart_post(sel_title = "Pull Factor Diffs in the Northeast and West", group = expost_region, facet = "expost_region", facet_titles = c("Northeast" = "Northeast", "West" = "West"))
northeast_west_post_diffs_chart

gc()
#Ante
# region_ante_trends_chart <- ante_migrant_trends_df %>%
#   group_by(year, exante_region, subgroup) %>%
#   summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
#   select(-frac_migrate_se) %>%
#   ungroup %>%
#   make_trend_chart(sel_title = "Push Factor Trends by Region", facet = "exante_region", facet_titles = c("Midwest" = "Midwest", "West" = "West", "South" = "South", "Northeast" = "Northeast")) #use male/female?

gc()
region_ante_diffs_df <- post_migrant_trends_df %>%
  group_by(year, exante_region, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se) %>%
  ungroup %>%
  pivot_wider(names_from = subgroup, values_from = frac_migrate) %>%
  mutate(samesex_diff = samesex_oldlegal - samesex_newlegal) %>%
  mutate(oppositesex_diff = oppositesex_oldlegal - oppositesex_newlegal) 

midwest_south_ante_diffs_chart <- region_ante_diffs_df %>%
  filter(exante_region %in% c("Midwest", "South")) %>%
  facet_make_diffs_chart_ante(sel_title = "Push Factor Diffs in the Midwest and South", group = exante_region, facet = "exante_region", facet_titles = c("Midwest" = "Midwest", "South" = "South"))
midwest_south_ante_diffs_chart

northeast_west_ante_diffs_chart <- region_ante_diffs_df %>%
  filter(exante_region %in% c("Northeast", "West")) %>%
  facet_make_diffs_chart_ante(sel_title = "Push Factor Diffs in the Northeast and West", group = exante_region, facet = "exante_region", facet_titles = c("Northeast" = "Northeast", "West" = "West"))
northeast_west_ante_diffs_chart
#takeaways: South/Midwest very interesting, some logic issues, surprisng stability/clean trends
#watch missing issues

#Race

gc()
###Export (oi map)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/sex_post_trends.png", plot = sex_post_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/men_post_diffs.png", plot = men_post_diffs_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/women_post_diffs.png", plot = women_post_diffs_chart)

# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/sex_ante_trends.png", plot = sex_ante_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/men_ante_diffs.png", plot = men_ante_diffs_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/women_ante_diffs.png", plot = women_ante_diffs_chart)

#consider these later
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/age_post_trends.png", plot = age_post_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/age_post_diffs.png", plot = age_post_diffs_chart)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/age_ante_trends.png", plot = age_ante_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/age_ante_diffs.png", plot = age_ante_diffs_chart)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/educ_post_trends.png", plot = educ_post_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/educ_post_diffs.png", plot = educ_post_diffs_chart)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/educ_ante_trends.png", plot = educ_ante_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/educ_ante_diffs.png", plot = educ_ante_diffs_chart)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/child_post_trends.png", plot = child_post_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/child_post_diffs.png", plot = child_post_diffs_chart)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/child_ante_trends.png", plot = child_ante_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/child_ante_diffs.png", plot = child_ante_diffs_chart)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/inc_post_trends.png", plot = inc_post_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/inc_post_diffs.png", plot = inc_post_diffs_chart)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/inc_ante_trends.png", plot = inc_ante_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/inc_ante_diffs.png", plot = inc_ante_diffs_chart)


# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/region_post_trends.png", plot = region_post_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/region_post_diffs.png", plot = midwest_south_post_diffs_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/region_post_diffs_app.png", plot = northeast_west_post_diffs_chart)
# ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/region_ante_trends.png", plot = region_ante_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/region_ante_diffs.png", plot = midwest_south_ante_diffs_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/region_ante_diffs_app.png", plot = northeast_west_ante_diffs_chart)
#more slicing and dicing to do

#more functions?
#separate all collapsing so runs better?
#watch more consistent labeling
#don't do trends just do differences?
#oop make things more efficient so don't have to summarize 4 times (and only twice)
#hmm wonder how much fixed effects would absorbr
#can totally functionalize/map all of this
#for triple diff charts makes sense to add another layer of difference?
#watch rename axes (Not/Federally)
#really wonder how much is motivated by noise/migration levels overall
#only look at region ante? still closed loop issue
```

MA chart (is this set up right? keep thinking)
```{r, include = F}
#CHECK TYPOS
gc()
#Pull Factor
#df
MA_post_migrant_trends_df <- MA_summary_stats %>%
  filter(expost_state == "MA") %>%
  mutate(subgroup = case_when(
    in_samesex == 1 ~ "samesex",
    in_samesex == 0 ~ "oppositesex"
  )) %>%
  group_by(year, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se)

#trends chart for export
gc()
MA_post_migrant_trends_chart <- MA_post_migrant_trends_df %>%
  ggplot() +
    geom_line(aes(x = year, y = frac_migrate, linetype = subgroup)) + 
    labs(title = "Pull Factor MA Trends Visualized",
         x = "Year",
         y = "Fraction Migrated",
         linetype = "Subgroup") + #maybe move title to plot itself
    geom_vline(xintercept = 2003, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2008, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2009, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2011, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2012, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2013, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2014, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2015, color = "black", linetype = "dashed") + 
    scale_linetype_manual(values = c("samesex" = "solid", "oppositesex" = "dotted"),
                     labels = c("samesex" = "Same-Sex Individuals", "oppositesex" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2001:2019) +
  theme_apa() #can always go back to theme

MA_post_migrant_trends_chart

#ante
#df
MA_ante_migrant_trends_df <- MA_summary_stats %>%
  filter(exante_state == "MA") %>%
  mutate(subgroup = case_when(
    in_samesex == 1 ~ "samesex",
    in_samesex == 0 ~ "oppositesex"
  )) %>%
  group_by(year, subgroup) %>%
  summarize(frac_migrate = survey_mean(migrant, na.rm = T)) %>%
  select(-frac_migrate_se)


#trends chart for export
gc()
MA_ante_migrant_trends_chart <- MA_ante_migrant_trends_df %>%
  ggplot() +
    geom_line(aes(x = year, y = frac_migrate, linetype = subgroup)) + 
    labs(title = "Push Factor MA Trends Visualized",
         x = "Year",
         y = "Fraction Migrated",
         linetype = "Subgroup") + #maybe move title to plot itself
    geom_vline(xintercept = 2003, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2008, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2009, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2011, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2012, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2013, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2014, color = "black", linetype = "dashed") + 
    geom_vline(xintercept = 2015, color = "black", linetype = "dashed") + 
    scale_linetype_manual(values = c("samesex" = "solid", "oppositesex" = "dotted"),
                     labels = c("samesex" = "Same-Sex Individuals", "oppositesex" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2001:2019) +
  theme_apa() #can always go back to theme

MA_ante_migrant_trends_chart

#export charts
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/MA_post_trends.png", plot = MA_post_migrant_trends_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/MA_ante_trends.png", plot = MA_ante_migrant_trends_chart)

#make difference tables?
#maybe
#lol event study uses, note vertical lines are time periods other states legalized
```

Predicted Charts (for illustration purposes)
```{r, include = F}
#Function:
make_ex_diffs_chart_post <- function(df, sel_title) {
  chart <- df %>%
  ggplot() +
    geom_line(aes(x = year, y = diffs, linetype = subgroup)) + 
    labs(title = sel_title,
         x = "Year",
         y = "Relative Migration to Obergefell States",
         linetype = "Subgroup") + 
    geom_vline(xintercept = 2015, color = "black", linetype = "dotted") + 
    scale_linetype_manual(values = c("same_sex" = "solid", "opposite_sex" = "longdash"), #adjust this as needs be later
                     labels = c("same_sex" = "Same-Sex Individuals", "opposite_sex" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa() +
  theme(axis.text.y = element_blank())
  return(chart)
}
make_ex_diffs_chart_ante <- function(df, sel_title) {
  chart <- df %>%
  ggplot() +
    geom_line(aes(x = year, y = diffs, linetype = subgroup)) + 
    labs(title = sel_title,
         x = "Year",
         y = "Relative Migration from Obergefell States",
         linetype = "Subgroup") + 
    geom_vline(xintercept = 2015, color = "black", linetype = "dotted") + 
    scale_linetype_manual(values = c("same_sex" = "solid", "opposite_sex" = "longdash"), #adjust this as needs be later
                     labels = c("same_sex" = "Same-Sex Individuals", "opposite_sex" = "Opposite-Sex Individuals")) +
  scale_x_continuous(breaks = 2011:2019) +
  theme_apa() +
  theme(axis.text.y = element_blank())
  return(chart)
}


#Make predicted dfs/charts
post_ex <- data.frame(
  year = 2011:2019,
  same_sex = c(-.5, -.5, -.5, -.5, -.2, -.2, -.2, -.2, -.2),
  opposite_sex = c(0, 0, 0, 0, 0, 0, 0, 0, 0)
  ) %>%
  pivot_longer(same_sex:opposite_sex, names_to = "subgroup", values_to = "diffs")

ante_ex <- data.frame(
  year = 2011:2019,
  same_sex = c(.5, .5, .5, .5, .2, .2, .2, .2, .2),
  opposite_sex = c(0, 0, 0, 0, 0, 0, 0, 0, 0)
) %>%
  pivot_longer(same_sex:opposite_sex, names_to = "subgroup", values_to = "diffs")

#Make charts
ex_post_diffs_chart <- post_ex %>%
  make_ex_diffs_chart_post(sel_title = "Predicted Pull Factor Differences")

ex_ante_diffs_chart <- ante_ex %>%
  make_ex_diffs_chart_ante(sel_title = "Predicted Push Factor Differences")

#export charts
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/ex_post_diffs.png", plot = ex_post_diffs_chart)
ggsave("/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/ex_ante_diffs.png", plot = ex_ante_diffs_chart)


#could write out better/separate functions out clearer
#output all together one line?
```

Code
```{r, include = F}
gc()
#code
```

Pop Support Table
```{r, include = F}
gc()
pop_support <- clean_df %>% #don't need survey object here
  distinct(expost_state, expost_support) %>%
  mutate(support = 100 - expost_support) %>% #need to update this
  arrange(support) %>%
  rename("State" = expost_state, "Support Level" = "support") %>%
  select(State, "Support Level") 

pop_support_export <-pop_support %>%
  kable(caption = "Support for Gay Men and Lesbian Women by State in 2016", format = "latex", booktabs = F, digits = 0, longtable = T)

writeLines(pop_support_export, "/Users/njrich/Desktop/same-sex-migration/outputs/summary_stats/pop_support.tex")

```

Test
```{r, include = F}
#check coding makes sense
test <- clean_df %>%
  select(year, migrant, exante_state, exante_old_legal, expost_state, expost_old_legal) %>%
  distinct() %>%
  filter(expost_state != exante_state) 

test2 <- clean_df %>%
  select(year, exante_state, exante_old_legal) %>%
  filter(exante_old_legal == 1) %>%
  distinct()

#seems to be good?
```