\documentclass[12pt,letterpaper]{article}
\usepackage{datetime} %maybe don't need
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{url}
\usepackage{setspace}
\usepackage{geometry}
\usepackage{graphicx} %I added this, we'll see

% Adjust margins to fit the AER style
\geometry{top=1in, bottom=1in, left=1in, right=1in}

% Title setup
\title{Title of the Paper}
\author{Noah Rich\footnote{University of Michigan, njrich@umich.edu.}}
\date{\today}

% Start the document
\begin{document}

\maketitle
\footnote{Acknowledgements:}

% Abstract section
\begin{abstract}
    Your abstract goes here. Keep it concise, summarizing the key contributions of the paper in no more than 150 words.
\end{abstract}

\newpage

% Main content section
\section{Introduction}

Write the introduction here, explaining the problem, significance, and structure of the paper.

\section{Literature Review}
Discuss relevant literature and establish the gap your paper addresses.

\section{Model Setup or Theoretical Framework}
Explain any models, theories, or key ideas here. You can use equations like the following example:

\begin{equation}
    y = \beta_0 + \beta_1 x + \epsilon
\end{equation}

\section{Empirical Analysis}
Provide any empirical analysis or methodology. Use tables and figures here, for example:

\begin{table}[h!]
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        Variable & Coefficient & Std. Error \\
        \hline
        $x_1$ & 1.23 & 0.45 \\
        $x_2$ & 0.67 & 0.12 \\
        \hline
    \end{tabular}
    \caption{Regression Results}
\end{table}

\section{Results}
Present and interpret the results. Include tables, graphs, or any statistical outputs.

\section{Conclusion}
Summarize the findings, implications, and suggest possible directions for future research.

\newpage

% References section
\begin{thebibliography}{99}

\bibitem{key1}
Author, A. (Year). \textit{Title of the Book or Paper}. Journal Name, Volume(Issue), Pages.

\bibitem{key2}
Author, B., & Author, C. (Year). \textit{Title of Another Paper}. Another Journal, Volume(Issue), Pages.

\end{thebibliography}

\end{document}